### 2020-07-17 Model_Build.Rmd
Cleaned datasets to build model

---
title: "Model Building"
author: "Jake Esprabens"
date: "7/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fable)
library(lubridate)
library(tsibble)
library(feasts)
library(zoo)
```

## Loading Data
```{r}
gt <- read.csv("MLB_Google_Trends.csv")
ar <- read.csv("MLB_Attendance_and_Revenue.csv")
po <- read.csv("MLB_Preseason_Odds.csv")
tp <- read.csv("MLB_Team_Performance.csv")
tms <- read.csv("MLB_TV_Market_Size.csv")
```

## Cleaning Google Trends Data
```{r}
# cleaning google trends data by replacing <1 with 0 and making hits numeric
# also converting to tsibble
clean_gt <- gt %>%
  mutate(date = as.character(date),
         year = as.character(year(date)),
         hits = str_replace(hits, "<1","0"),
         hits = as.numeric(hits)) %>%
  rename(team = teams) %>%
  select(-X)
```

## Cleaning Preseason Odds Data
```{r}
clean_po <- po %>%
  pivot_longer(cols = 2:18, names_to = "year") %>%
  mutate(year = substring(year, 2)) %>%
  rename(team = Team,
         odds = value)
```

## Cleaning TV Market Size
```{r}
clean_tms <- tms %>%
  mutate_all(na_if,"") %>%
  mutate(TVMarketSize = as.numeric(gsub(",","",TVMarketSize))) %>%
  rename(team = Team)
```

## Cleaning Attendance and Revenue dataset 
```{r}
clean_ar <- ar %>%
  select(Year, Tm, Attendance, Attend.G, Est..Payroll) %>%
  mutate(Attendance = as.numeric(gsub(",","",Attendance)),
         Attend.G = as.numeric(gsub(",","",Attend.G)),
         Est..Payroll = as.numeric(str_replace_all(Est..Payroll, "[^[:alnum:]]", ""))) %>%
  rename(year = Year,
         attendance = Attendance,
         att_per_game = Attend.G,
         payroll = Est..Payroll)
```

## Cleaning Team Performance
```{r}
clean_tp <- tp %>%
  mutate(team = name, 
         year = yearID,
         team = (gsub("Anaheim Angels","Los Angeles Angels of Anaheim", team)),
         team = (gsub("Florida Marlins","Miami Marlins", team))) %>%
  select(year, team, W, L, DivWin, WCWin, LgWin, WSWin, attendance) %>%
  arrange(year, team)
```

## Combining Datasets
```{r}
combined <- left_join(clean_gt, clean_po, by = c("team", "year")) %>%
  mutate(date = yearmonth(date)) %>%
  select(-year) %>%
  as_tsibble(key = team, index = date)
```

# Building Model
```{r}
model_test <- combined %>%
  filter(team =="Chicago Cubs") %>%
  model(
    snaive = SNAIVE(hits),
    ets = ETS(hits),
    arima = ARIMA(hits)
  )
```

### 2020-07-20 Model_Build.Rmd
Touched up datasets

@@
   rename(team = teams) %>%
-  select(-X)
+  select(-X) %>%
+  arrange(team)
 ```
@@
 ```
-
-## Cleaning Attendance and Revenue dataset 
-```{r}
-clean_ar <- ar %>%
-  select(Year, Tm, Attendance, Attend.G, Est..Payroll) %>%
-  mutate(Attendance = as.numeric(gsub(",","",Attendance)),
-         Attend.G = as.numeric(gsub(",","",Attend.G)),
-         Est..Payroll = as.numeric(str_replace_all(Est..Payroll, "[^[:alnum:]]", ""))) %>%
-  rename(year = Year,
-         attendance = Attendance,
-         att_per_game = Attend.G,
-         payroll = Est..Payroll)
-```
-
 ## Cleaning Team Performance
@@
 
+## Cleaning Attendance and Revenue dataset 
+```{r}
+clean_ar <- ar %>%
+  mutate(attendance = as.numeric(gsub(",","",Attendance)),
+         att_per_game = as.numeric(gsub(",","",Attend.G)),
+         payroll = as.numeric(str_replace_all(Est..Payroll, "[^[:alnum:]]", ""))) %>%
+  rename(year = Year) %>%
+  select(year, Tm, attendance, att_per_game, payroll)
+
+clean_ar <- left_join(clean_ar, clean_tp, by = c("year", "attendance")) %>%
+  select(year, team, attendance, att_per_game, payroll)
+```
+
 ## Combining Datasets
@@
   as_tsibble(key = team, index = date)
+
+combined
 ```
@@
   )
+
+model_test$arima
 ```

### 2020-07-21 Model_Build.Rmd
Fixed proportions on TVMarketSize

@@
          hits = as.numeric(hits)) %>%
-  rename(team = teams) %>%
+  rename(team = teams,
+         googlehits = hits) %>%
   select(-X) %>%
@@
   rename(team = Team)
+
+# CWS/CHC, LAA/LAD, NYM/NYY, OAK/SFG all share TV Markets so must proportion by TV Rating
+clean_tms <- clean_tms %>%
+  mutate(proportions = case_when(
+    team == "Chicago Cubs" ~ (4.43 / (4.43 + 1.03)),
+    team == "Chicago White Sox" ~ (1.03 / (4.43 + 1.03)),
+    team == "Los Angeles Angels" ~ (1.43 / (1.43 + 1.96)),
+    team == "Los Angeles Dodgers" ~ (1.96 / (1.43 + 1.96)),
+    team == "New York Mets" ~ (2.47 / (2.47 + 3.34)),
+    team == "New York Yankees" ~ (3.34 / (2.47 + 3.34)),
+    team == "Oakland Athletics" ~ (0.88 / (0.88 + 2.68)),
+    team == "San Francisco Giants" ~ (0.88 / (0.88 + 2.68)),
+    TRUE ~ 1
+    )) %>%
+  mutate(propTVMarketSize = round(TVMarketSize * proportions)) %>%
+  select(team, propTVMarketSize)
 ```
+
 ## Cleaning Team Performance
@@
   model(
-    snaive = SNAIVE(hits),
-    ets = ETS(hits),
-    arima = ARIMA(hits)
+    snaive = SNAIVE(googlehits),
+    ets = ETS(googlehits),
+    arima = ARIMA(googlehits)
   )

### 2020-07-22 Model_Build.Rmd
Cleaned data again

--- 2020-07-17 Model_Build.Rmd
@@
 ```{r}
-# cleaning google trends data by replacing <1 with 0 and making hits numeric
-# also converting to tsibble
+# cleaning google trends data by replacing <1 with 0 and and changing type of other variables
 clean_gt <- gt %>%
@@
          hits = as.numeric(hits)) %>%
-  rename(team = teams) %>%
-  select(-X)
+  rename(team = teams,
+         googlehits = hits) %>%
+  select(-X) %>%
+  arrange(team)
 ```
@@
 ```{r}
+# changed the structure of the data using pivot_longer and then changed certain names of columns
 clean_po <- po %>%
@@
 ```{r}
+# take out all commas in observations
 clean_tms <- tms %>%
@@
   rename(team = Team)
+
+# CWS/CHC, LAA/LAD, NYM/NYY, OAK/SFG all share TV Markets so must proportion by TV Rating to split the Market size
+clean_tms <- clean_tms %>%
+  mutate(proportions = case_when(
+    team == "Chicago Cubs" ~ (4.43 / (4.43 + 1.03)),
+    team == "Chicago White Sox" ~ (1.03 / (4.43 + 1.03)),
+    team == "Los Angeles Angels" ~ (1.43 / (1.43 + 1.96)),
+    team == "Los Angeles Dodgers" ~ (1.96 / (1.43 + 1.96)),
+    team == "New York Mets" ~ (2.47 / (2.47 + 3.34)),
+    team == "New York Yankees" ~ (3.34 / (2.47 + 3.34)),
+    team == "Oakland Athletics" ~ (0.88 / (0.88 + 2.68)),
+    team == "San Francisco Giants" ~ (2.68 / (0.88 + 2.68)),
+    TRUE ~ 1
+    )) %>%
+  mutate(propTVMarketSize = round(TVMarketSize * proportions)) %>%
+  select(team, propTVMarketSize)
+```
+
+## Cleaning Team Performance
+```{r}
+# changed Angels, Marlins, and Expos to all their updated names / cities
+# changed the N and Y to 0's and 1's to indicate whether each team won their division, wild card game, league, and world series
+clean_tp <- tp %>%
+  mutate(team = name, 
+         year = yearID,
+         team = (gsub("Anaheim Angels","Los Angeles Angels of Anaheim", team)),
+         team = (gsub("Florida Marlins","Miami Marlins", team)),
+         team = (gsub("Montreal Expos","Washington Nationals", team))) %>%
+  mutate(year = as.character(year),
+         team = as.factor(team)) %>%
+  mutate(DivWin = as.factor(case_when(DivWin == "N" ~ 0, DivWin == "Y" ~ 1)),
+         WCWin = as.factor(case_when(WCWin == "N" ~ 0, WCWin == "Y" ~ 1)),
+         LgWin = as.factor(case_when(LgWin == "N" ~ 0, LgWin == "Y" ~ 1)),
+         WSWin = as.factor(case_when(WSWin == "N" ~ 0, WSWin == "Y" ~ 1))) %>%
+  select(year, team, W, L, DivWin, WCWin, LgWin, WSWin, attendance) %>%
+  arrange(year, team)
 ```
@@
 ```{r}
+# took out commas in all numeric observations
 clean_ar <- ar %>%
-  select(Year, Tm, Attendance, Attend.G, Est..Payroll) %>%
-  mutate(Attendance = as.numeric(gsub(",","",Attendance)),
-         Attend.G = as.numeric(gsub(",","",Attend.G)),
-         Est..Payroll = as.numeric(str_replace_all(Est..Payroll, "[^[:alnum:]]", ""))) %>%
-  rename(year = Year,
-         attendance = Attendance,
-         att_per_game = Attend.G,
-         payroll = Est..Payroll)
-```
+  rename(year = Year) %>%
+  mutate(attendance = as.numeric(gsub(",","",Attendance)),
+         att_per_game = as.numeric(gsub(",","",Attend.G)),
+         payroll = as.numeric(str_replace_all(Est..Payroll, "[^[:alnum:]]", "")),
+         year = as.character(year),
+         team = as.factor(Tm)) %>%
+  select(year, Tm, attendance, att_per_game, payroll)
 
-## Cleaning Team Performance
-```{r}
-clean_tp <- tp %>%
-  mutate(team = name, 
-         year = yearID,
-         team = (gsub("Anaheim Angels","Los Angeles Angels of Anaheim", team)),
-         team = (gsub("Florida Marlins","Miami Marlins", team))) %>%
-  select(year, team, W, L, DivWin, WCWin, LgWin, WSWin, attendance) %>%
-  arrange(year, team)
+# needed to get full names of teams so used team names from previous dataframe
+clean_ar <- left_join(clean_ar, clean_tp, by = c("year", "attendance")) %>%
+  select(year, team, attendance, att_per_game, payroll)
 ```
@@
 ```{r}
-combined <- left_join(clean_gt, clean_po, by = c("team", "year")) %>%
+# combining all dataframes together
+combined <- left_join(clean_gt, clean_po, by = c("team", "year")) # combine google trends and preseason odds data
+combined <- left_join(combined, clean_tms, by = "team") # combine previous dataframe with TVMarketSize data
+combined <- left_join(combined, clean_tp, by = c("team", "year")) # combine previous dataframe with team performance data
+combined <- left_join(combined, clean_ar, by = c("team", "year")) # combine previous dataframe with attendance and revenue data
+
+# picking out and renaming specific variables
+# converting date to a month object to fit in fable / tsibble functions
+# convert to tsibble
+alldata <- combined %>%
+  rename(attendance = attendance.x) %>%
+  select(-c(year, attendance.y)) %>%
   mutate(date = yearmonth(date)) %>%
-  select(-year) %>%
   as_tsibble(key = team, index = date)
@@
 ```{r}
-model_test <- combined %>%
+model_test <- alldata %>%
   filter(team =="Chicago Cubs") %>%
   model(
-    snaive = SNAIVE(hits),
-    ets = ETS(hits),
-    arima = ARIMA(hits)
+    arima = ARIMA(googlehits)
   )
+
+model_test$arima
 ```

### 2020-07-27 Model_Build.Rmd
Updated Model and Started Write up

@@
 library(feasts)
-library(zoo)
 ```
@@
          WSWin = as.factor(case_when(WSWin == "N" ~ 0, WSWin == "Y" ~ 1))) %>%
-  select(year, team, W, L, DivWin, WCWin, LgWin, WSWin, attendance) %>%
+  mutate(winpct = round(W / (W + L), digits = 3)) %>%
+  select(year, team, winpct, DivWin, WCWin, LgWin, WSWin, attendance) %>%
   arrange(year, team)
@@
   mutate(attendance = as.numeric(gsub(",","",Attendance)),
-         att_per_game = as.numeric(gsub(",","",Attend.G)),
+         attgame = as.numeric(gsub(",","",Attend.G)),
          payroll = as.numeric(str_replace_all(Est..Payroll, "[^[:alnum:]]", "")),
@@
          team = as.factor(Tm)) %>%
-  select(year, Tm, attendance, att_per_game, payroll)
+  select(year, Tm, attendance, attgame, payroll)
 
@@
 clean_ar <- left_join(clean_ar, clean_tp, by = c("year", "attendance")) %>%
-  select(year, team, attendance, att_per_game, payroll)
+  select(year, team, attendance, attgame, payroll)
 ```
@@
   mutate(date = yearmonth(date)) %>%
-  as_tsibble(key = team, index = date)
+  as_tsibble(key = c(team, propTVMarketSize, odds, winpct), index = date)
 ```
@@
 ```{r}
-model_test <- alldata %>%
-  filter(team =="Chicago Cubs") %>%
+# filtering down to one team
+lad <- alldata %>%
+  filter(team == "Los Angeles Dodgers") %>%
+  arrange(date)
+
+# creating snaive, ets, and arima models
+fit <- lad %>%
   model(
-    arima = ARIMA(googlehits)
+    arima = ARIMA(googlehits) # testing with only arima right now
   )
 
-model_test$arima
+# shows coefficients of the arima models
+# Problem with have multiple coefficients
+# PROBLEM?? Confused becauses doesnt account for every date so it only creates estimates and models for each unique row (per season)
+
+coef <- fit %>%
+  select(team, arima) %>%
+  coef()
+
+#forecasted 12 periods into future / do i need to forecast?
+# fc <- fit %>%
+#   forecast(h = 12)
+# 
+# fc %>%
+#   autoplot(lad, level = NULL)
 ```

### 2020-07-27 Write_Up.Rmd
Updated Model and Started Write up

---
title: "MLB Write Up"
author: "Jake Esprabens"
date: "7/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Collection

In order to measure the bandwagon level of MLB fans, we gathered data from google trends to depict the popularity of each team.  This data was gathered using the **gtrendsR** package, which automatically pulls the data from Google Trends API.  The units of the Google Trend data are relative to the maximum

### 2020-07-27 Write_Up.Rmd
Added to Write up and model

@@
 In order to measure the bandwagon level of MLB fans, we gathered data from google trends to depict the popularity of each team.
-This data was gathered using the **gtrendsR** package, which automatically pulls the data from Google Trends API.
-The units of the Google Trend data are relative to the maximum
+This data was gathered using the [**gtrendsR**](https://cran.r-project.org/web/packages/gtrendsR/gtrendsR.pdf) package, which automatically pulls the data from [Google Trends](https://trends.google.com/trends/?geo=US) API.
+The units of the Google Trend data are relative to the maximum of the trend.
+Therefore, we went through each team and determined that the Chicago Cubs had the maximum google trend hits; this lead us to gather data for the rest of the MLB teams relative to the Cubs.
+
+The preseason odds dataset was gathered using data from [Sports Odds History](https://www.sportsoddshistory.com/mlb-odds/).
+We compiled this data by viewing each teams' odds of winning the World Series from 2004 to the present.
+Ideally, we would have like to use monthly odds of winning the World Series for each MLB team; however, this data was not available for every desired year.

### 2020-07-30 Model_Build.Rmd
Added to Write up and model
Added plot of fitted time series

@@
 library(feasts)
+library(mable)
 ```
@@
   mutate(date = yearmonth(date)) %>%
-  as_tsibble(key = c(team, propTVMarketSize, odds, winpct), index = date)
+  as_tsibble(key = team, index = date)
 ```
@@
   model(
-    arima = ARIMA(googlehits) # testing with only arima right now
+    arima = ARIMA(googlehits ~ odds + attgame + winpct) # testing with only arima right now
   )
 
-# shows coefficients of the arima models
-# Problem with have multiple coefficients
-# PROBLEM?? Confused becauses doesnt account for every date so it only creates estimates and models for each unique row (per season)
+# ARIMA is used when implementing exogenous regressors
+
+# DIFFERENT FUNCTIONS TO USE ON MABLE
+# augment() grabs fitted values and we can plot these
+# glance() checks AIC and BIC
+# components() allows us to use ggplot to plot aspects of the model / only works on ets and snaive i think
+
+fitted <- augment(fit)
+
+fitted %>%
+  ggplot() + 
+  geom_line(aes(x = date, y = googlehits), color = "blue") +
+  geom_line(aes(x = date, y = .fitted), color = "red")
+
+# small problem is that it can't produce fitted values for any 2020 months because I don't have winpct for those dates / dont know if it matters?
 
@@
 
-#forecasted 12 periods into future / do i need to forecast?
-# fc <- fit %>%
-#   forecast(h = 12)
-# 
-# fc %>%
-#   autoplot(lad, level = NULL)
+coef
 ```

### 2020-07-30 Model_Build.Rmd
Added comments around model building

@@
 
-# ARIMA is used when implementing exogenous regressors
+# ARIMA is used when implementing exogenous regressors so Arima works for our data
+# from what I have seen, the date doesn't need to me on the right hand side of the formula
+# I believe you can manually add seasonality if you specify your own model; however, it should estimate seasonality if I don't specify it
 
@@
 
+#PLOT of Actual vs Fitted Values
 fitted %>%
@@
 coef
+# exogenous regressors are still not significant though?
 ```

### 2020-08-03 Model_Build.Rmd
Added interpretations for arima model and added hierarchical model

@@
 
-# creating snaive, ets, and arima models
+# creating arima models
 fit <- lad %>%
   model(
-    arima = ARIMA(googlehits ~ odds + attgame + winpct) # testing with only arima right now
+    arima = ARIMA(googlehits ~ odds + attgame + winpct + payroll + DivWin + LgWin + WSWin) # testing with only some regressors
   )
@@
 ```
+
+## Interpretting Time Series Terms
+### Dodger Example
+
+ARIMA(p,d,q)(P,D,Q)s
+
+In this case focusing on just the Dodgers, we have an $ARIMA(0,0,0)(1,0,1)12$ model.
+The $12$ represents the numer of periods within a year; therefore, this is a monthly series.
+The seasonal autoregressive order 1 (sar1) term means that we are using a linear combination with the values of the variable of interest, google hits, lagged from exactly 12 months ago.
+The seasonal moving average order 1 (sma1) means that the current google hits are affected by some random error from exactly 12 months ago.
+
+The model without the exogenous regressors can be written as by $(Y_t-\phi_{12} Y_{t - 12}) = (e_{t}+\Theta_{12} e_{t-12})$  where $\phi_{12} = 0.509$ and $\Theta_{12} = 0.308$.
+This $0.308$ of the sar1 term can interpretted as moving $60.2\%$ of the way from the current google hits toward the average of the google hits.
+The $0.509$ of the
+
+I'm iffy on the numerical interpretations of these terms ^.
+Most people say that they are just different weights and either converge or diverge faster or slower depending on the estimate of the coefficient.
+
+## Hierarchial Model
+```{r}
+# all I did was add propTVMarketSize as a key when converting to tsibble, not sure if that's the correct way to do it?
+I didn't add propTVMarketSize to the arima model.
+It says it wasn't found if I add it there.
+
+alldata_hier <- combined %>%
+  rename(attendance = attendance.x) %>%
+  select(-c(year, attendance.y)) %>%
+  mutate(date = yearmonth(date)) %>%
+  as_tsibble(key = c(team, propTVMarketSize), index = date)
+
+hier_model <- alldata_hier %>%
+  model(
+    arima = ARIMA(googlehits ~ odds + attgame + winpct + payroll + DivWin + LgWin + WSWin)
+  )
+
+hier_fitted <- augment(hier_model)
+
+hier_model %>%
+  select(team, arima) %>%
+  coef()
+```

### 2020-08-04 Model_Build.Rmd
Fixed small problem with hier model

@@
          year = yearID,
-         team = (gsub("Anaheim Angels","Los Angeles Angels of Anaheim", team)),
+         team = (gsub("Anaheim Angels","Los Angeles Angels", team)),
+         team = (gsub("Los Angeles Angels of Anaheim","Los Angeles Angels", team)),
          team = (gsub("Florida Marlins","Miami Marlins", team)),
@@
   model(
-    arima = ARIMA(googlehits ~ odds + attgame + winpct + payroll + DivWin + LgWin + WSWin)
+    arima = ARIMA(googlehits ~ odds + attgame + winpct + payroll + DivWin + WCWin + LgWin + WSWin)
   )

### 2020-08-10 Model_Build.Rmd
Updated Write Up

@@
 
-The model without the exogenous regressors can be written as by $(Y_t-\phi_{12} Y_{t - 12}) = (e_{t}+\Theta_{12} e_{t-12})$  where $\phi_{12} = 0.509$ and $\Theta_{12} = 0.308$.
-This $0.308$ of the sar1 term can interpretted as moving $60.2\%$ of the way from the current google hits toward the average of the google hits.
-The $0.509$ of the
+The model without the exogenous regressors can be written as by $(Y_t-\phi_{12} Y_{t - 12}) = (e^{Yt}+\Theta_{12} e^{Yt-12})$  where $\phi_{12} = 0.509$ and $\Theta_{12} = 0.308$.
+This $0.308$ of the sar1 term can interpretted as moving $69.2\%$ of the way from the current google hits toward the average of the google hits.
 

### 2020-08-10 Write_Up.Rmd
Updated Write Up

@@
 
+### Google hits
 In order to measure the bandwagon level of MLB fans, we gathered data from google trends to depict the popularity of each team.
@@
 
+### Sports Betting Odds
 The preseason odds dataset was gathered using data from [Sports Odds History](https://www.sportsoddshistory.com/mlb-odds/).
@@
 Ideally, we would have like to use monthly odds of winning the World Series for each MLB team; however, this data was not available for every desired year.
+
+### Attendance and Payroll
+We decided that attendence and payroll of each team by season may be valuable exogenous variables that could help determine when fans would become bandwagons.
+We gathered this data from [Baseball Reference](https://www.baseball-reference.com/leagues/MLB/2019-misc.shtml) by copying the data from each the years 2004 to 2019.
+
+### Team Performance
+Other exogenous variables we decided would be a valuable predictor in this model was win percentage and whether each team made the playoffs.
+We were able to download this data directly from [Sean Lahman's Baseball Database](http://www.seanlahman.com/baseball-archive/statistics/).
+This dataset was also filled with other variables such as attendance along with indicators of whether each team won their respective divisions, wild card games, leagues, and the world series.
+
+### TV Market Size
+The TV Market Size of each team is was also important to create an accurate estimation of the number of true fans of each.
+We were able to gather this data using the Nielsen Market Size from [Sports Media Watch](https://www.sportsmediawatch.com/nba-market-size-nfl-mlb-nhl-nielsen-ratings/).

### 2020-08-11 KB_modeling.Rmd
still frustrate

---
title: "Hierarchical ARIMA"
author: "Kelly Bodwin"
date: "8/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fable)
library(lubridate)
library(tsibble)
library(feasts)
library(mable)
library(hts)
library(modelr)
```

```{r}
dat <- read_csv(here::here("combined_data.csv"))

dat <- dat %>%
  mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
  mutate(
    playoffs = case_when(
      WSWin ~ "Won WS",
      LgWin ~ "Won League",
      WCWin  ~ "Won Conference",
      DivWin  ~ "Won Division",
      TRUE ~ "Did not qualify"
    ),
    playoffs = forcats::fct_inorder(playoffs),
    phase = case_when(
      month < 04 | month > 10 ~ "Summer",
      month == 04 ~ "Opening",
      month < 10 & month > 04 ~ "Season",
      month == 10 ~ "Playoffs"
    ),
    phase = forcats::fct_reorder(phase, month),
    is_playoffs = as.integer(phase == "Playoffs")
  ) %>%
  mutate_at(vars(month, year, team), funs(factor))

summary(dat$playoffs)
```

**Forever** variables:
- `propTVMarketSize` = # of people in TV market region

**Start of Season** variables:
- `odds` = preseason odds of winning WS
- `payroll` = team's payroll at start of season (?)

**End of Season** variables:
- `attendance` = total attendance
- `attgame` = attendance per game (?)
- `winpct` = win percent of season
- `DivWin`, etc = dummy variables for playoff success

**Monthly** variables:
- Just `googlehits`

Overall mean google hits should be determined by market size.

Annual mean hits should be initially predicted by percieved competitiveness.

Annual mean hits should actually occur based on win percent and playoff success.

Annual mean hits are well measured by attendance.

Monthly hits should be determined by:
- Seasonality
- What happened last month with the team
- Playoff status

Do the "boosts" act as a percent of the possible volume?

---
## Explore data

```{r}
dat %>%
  ggplot() +
  geom_histogram(aes(x = log(googlehits))) +
  facet_wrap(~phase)

dat %>%
  filter(googlehits > 0) %>%
  summarize(mean(log(googlehits+1)))
```

Basically, ghits are very, very skewed, and this is a bit of a problem.

I'm going to do a silly cheat that is actually used a lot, and quantile normalize
this data.

```{r}
quant_norm <- function(vec){

  quants <- rank(vec)/(length(vec) + 1)
  normies <- purrr::map_dbl(quants, qnorm)

  return(normies)

}

dat <- dat %>%
  mutate(
    qn_ghits = quant_norm(googlehits)
  )

```

```{r}
CA_teams <- c("San Francisco Giants", "Los Angeles Dodgers", "San Diego Padres", 
              "Oakland Athletics")

example_teams <- c("Chicago Cubs", "Arizona Diamondbacks", "Los Angeles Dodgers")

## Just makes things a little less extreme.

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = date, y = qn_ghits, color = team)) +
  geom_line()

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = date, y = googlehits)) +
  geom_line()
```

---

```{r}
dat %>%
  group_by(playoffs, month) %>% 
  summarize(googlehits = mean(googlehits)) %>%
  ungroup() %>%
 ggplot(aes(x = month, y = googlehits)) +
  geom_point() +
  facet_wrap(~playoffs)

dat %>%
  group_by(playoffs, month) %>% 
  summarize(googlehits = mean(qn_ghits)) %>%
  ungroup() %>%
 ggplot(aes(x = month, y = googlehits)) +
  geom_point() +
  facet_wrap(~playoffs)

```

---
## Ugly model

```{r}
mod <- lm(qn_ghits ~ propTVMarketSize  + odds + attgame + payroll + is_playoffs*winpct +  is_playoffs*playoffs, data = dat)

summary(mod)

dat %>%
  add_predictions(mod) %>%
  mutate(
    error = qn_ghits - pred
  ) %>%
  filter(team %in% c("Boston Red Sox", "New York Yankees", "Arizona Diamondbacks")) %>%
  #filter(team == "Boston Red Sox") %>%
  ggplot() +
   geom_line(aes(x = date, y = error, color = team))
  #geom_line(aes(x = date, y = qn_ghits), color = "cornflowerblue") +
  #geom_line(aes(x = date, y = pred), color = "deeppink")
```

---
## More exploring

```{r}
by_team <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team) %>%  
  summarize(qn_ghits = mean(qn_ghits),
            propTVMarketSize = max(propTVMarketSize))

by_year <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team, year) %>%  
  summarize(qn_ghits = median(qn_ghits),
            attendance = max(attendance),
            payroll = max(payroll),
            odds = max(odds),
            winpct = max(winpct)) %>%
  ungroup()

by_year %>%
  ggplot(aes(x = payroll, y = qn_ghits)) +
  geom_point()

library(GGally)

by_year %>%
  select(-team, -year) %>%
  ggpairs()
```
---

## Model 1: Overall mean

```{r}
by_team %>%
  filter(team %in% example_teams)

global_means <- lm(qn_ghits ~ propTVMarketSize, data = by_team)

summary(global_means)
```

Takeaway:  More hits in bigger markets, as expected.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(global_means, var = "ghits_p1") %>%
  mutate(
    ghits_e1 = qn_ghits - ghits_p1
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e1, color = team)) +
  geom_line()
```

... but not much evening out happening here.  Seems that the overall levels probably
have more to do with team success than anything else.

---

## Model 2: Annual means

```{r}
by_year <- 
  dat %>%
  filter(phase == "Season") %>%
  group_by(team, year) %>%  
  summarize(ghits_e1 = mean(ghits_e1),
            attendance = max(attendance),
            payroll = max(payroll),
            odds = max(odds),
            winpct = max(winpct)) %>%
  ungroup()

annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll, data = by_year) ## this should be an AR model

summary(annual_means)

```

Takeaway:  Payroll and winpct are a good predictor of annual ghits

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(annual_means, var = "ghits_p2") %>%
  mutate(
    ghits_e2 = qn_ghits - ghits_p1 - ghits_p2
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e2, color = team)) +
  geom_line()
```

Much more evened out for Dodgers and Dbacks.  Why not for Cubs?  Probably because
of the later increasingtrend they have.

---

## Model 3: Monthly

```{r}
final_model <- 
  dat  %>%
  lm(ghits_e2 ~ phase + is_playoffs:playoffs, data = .) ## needs to be AR model instead

summary(final_model)
```

Takeaway:  More hits in bigger markets, as expected.

```{r}
## make new column of errors
dat <- dat %>%
  modelr::add_predictions(final_model, var = "ghits_p3") %>%
  mutate(
    ghits_e3 = ghits_e2 - ghits_p3,
    final_preds = ghits_p1 + ghits_p2 + ghits_p3
  )

dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = date, y = ghits_e3, color = team)) +
  geom_line()
```

Error seems a lot more random, which is the goal!

```{r}
dat %>%
  filter(team %in% CA_teams) %>%
  ggplot(aes(x = date, y = ghits_e3, color = team)) +
  geom_line()
```
```{r}
dat %>%
  filter(team %in% example_teams) %>%
  ggplot(aes(x = ghits_e3)) +
  geom_histogram() +
  facet_wrap(~team)
```

```{r}
dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot(aes(x = qn_ghits, y = final_preds, color = phase)) + 
  geom_point()

dat %>%
  filter(team == "Chicago Cubs") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Arizona Diamondbacks") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")

dat %>%
  filter(team == "Houston Astros") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = final_preds), color = "cornflowerblue")
```

Seems to be underpredicting heavily when teams are bad.  Hmmmm.  Maybe the AR model will help that.

---

# Next: A non-ARIMA hierarchical model

```{r}
lump_mod <- lm(qn_ghits ~ payroll + (odds + winpct)*phase + is_playoffs:playoffs, data = dat)

summary(lump_mod)

dat <- dat %>%
  modelr::add_predictions(lump_mod, var = "lump_preds")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")
```
This is much better.

```{r}
library(lme4)

lump_mod <- lmer(qn_ghits ~ payroll + odds + winpct + (is_playoffs:playoffs + 1|team), data = dat)

summary(lump_mod)

dat <- dat %>%
  modelr::add_predictions(lmer, var = "lump_preds")

dat %>%
  filter(team == "Los Angeles Dodgers") %>%
  ggplot() + 
  geom_line(aes(x = date, y = qn_ghits), color = "deeppink") +
  geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")
```

### 2020-08-11 KB_modeling.Rmd
add some notes

@@
 summary(dat$playoffs)
+
+## playoffs variable
+#  Did Not make Playoffs or WC Play-in 
+#  Played in WC Play-In
+#  Eliminated Round 1
+#  Eliminated Round 2
+#  Lost WS
+#  Won WS
 ```
@@
   ggplot() +
-  geom_histogram(aes(x = log(googlehits))) +
-  facet_wrap(~phase)
+  geom_histogram(aes(x = log(googlehits))) #+
+  #facet_wrap(~phase)
 
@@
 dat %>%
-  filter(team == "Houston Astros") %>%
+  filter(team == "Baltimore Orioles") %>%
   ggplot() + 
@@
 ```{r}
-lump_mod <- lm(qn_ghits ~ payroll + (odds + winpct)*phase + is_playoffs:playoffs, data = dat)
+lump_mod <- lm(qn_ghits ~ payroll + (odds + winpct)*phase + is_playoffs:playoffs:team, data = dat)
 
@@
   geom_line(aes(x = date, y = lump_preds), color = "cornflowerblue")
-```
-This is much better.
-
-```{r}
-library(lme4)
-
-lump_mod <- lmer(qn_ghits ~ payroll + odds + winpct + (is_playoffs:playoffs + 1|team), data = dat)
-
-summary(lump_mod)
-
-dat <- dat %>%
-  modelr::add_predictions(lmer, var = "lump_preds")
 
 dat %>%
-  filter(team == "Los Angeles Dodgers") %>%
+  filter(team == "San Diego Padres") %>%
   ggplot() + 
@@
 ```
+This is much better.

### 2020-08-15 KB_modeling.Rmd
Added new dataset of playoff success

@@
 
+## Added new dataset for playoff
+playoff_success <- read.csv("MLB_Postseason_Success.csv")
+
 dat <- dat %>%
-  mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
+  # mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
   mutate(
-    playoffs = case_when(
-      WSWin ~ "Won WS",
-      LgWin ~ "Won League",
-      WCWin  ~ "Won Conference",
-      DivWin  ~ "Won Division",
-      TRUE ~ "Did not qualify"
-    ),
-    playoffs = forcats::fct_inorder(playoffs),
+  #   playoffs = case_when(
+  #     WSWin ~ "Won WS",
+  #     LgWin ~ "Lost WS",
+  #     WCWin  ~ "Won Wild Card",
+  #     DivWin  ~ "Won Division",
+  #     TRUE ~ "No Playoffs"
+  #   ),
+    # playoffs = forcats::fct_inorder(playoffs),
     phase = case_when(
@@
     phase = forcats::fct_reorder(phase, month),
-    is_playoffs = as.integer(phase == "Playoffs")
+    # is_playoffs = as.integer(phase == "Playoffs")
   ) %>%
@@
 #  Did Not make Playoffs or WC Play-in
-#  Played in WC Play-In
-#  Eliminated Round 1
-#  Eliminated Round 2
-#  Lost WS
-#  Won WS
+#  Played in WC Play-In - Going to need to look up and manually add (Is this only the team that lost the wild card?)
+#  Eliminated Round 1 - Going to need to look up and manually add
+#  Eliminated Round 2 - Going to need to look up and manually add
+#  Lost WS - LgWin but no WSwin
+#  Won WS - WSWin 
 ```

### 2020-08-15 Model_Build.Rmd
Added new dataset of playoff success

@@
 
-## Hierarchial Model
+## Hierarchial Model (Note Right)
 ```{r}

### 2020-08-16 KB_modeling.Rmd
Combined new playoff data and added ar terms

--- 2020-08-11 KB_modeling.Rmd
@@
 
+## Added new dataset for playoff
+# broken up into 6 different factors
+# - Won WS (1 team)
+# - Lost WS (1 team)
+# - Lost Semi (2 teams)
+# - Lost Quarter (4 teams)
+# - Lost WC (2 teams)    NOTE: wildcard game started in 2012 so this variable is only from 2012-2019.
+# - No Playoffs (22 teams from 2004-2011 and 20 teams from 2012-2019)
+
+playoff_success <- read.csv("MLB_Postseason_Success.csv", stringsAsFactors = FALSE) %>%
+  mutate_all(na_if,"") %>%
+  replace(is.na(.), "No Playoffs")
+
+#combining new palyoff dataset with dat
+dat <- left_join(dat, playoff_success, by = c("team", "year")) %>%
+  select(-c(DivWin, WCWin, LgWin, WSWin)) # deselecting unnecessary variables
+
 dat <- dat %>%
-  mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
+  # mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
   mutate(
-    playoffs = case_when(
-      WSWin ~ "Won WS",
-      LgWin ~ "Won League",
-      WCWin  ~ "Won Conference",
-      DivWin  ~ "Won Division",
-      TRUE ~ "Did not qualify"
-    ),
-    playoffs = forcats::fct_inorder(playoffs),
+  #   playoffs = case_when(
+  #     WSWin ~ "Won WS",
+  #     LgWin ~ "Lost WS",
+  #     WCWin  ~ "Won Wild Card",
+  #     DivWin  ~ "Won Division",
+  #     TRUE ~ "No Playoffs"
+  #   ),
+    playoffs = forcats::fct_inorder(playoffs), # not exactly sure what this line of code does? do we still need?
     phase = case_when(
@@
     ),
-    phase = forcats::fct_reorder(phase, month),
+    phase = forcats::fct_reorder(phase, month), # not exactly sure what this line of code does? do we still need?
     is_playoffs = as.integer(phase == "Playoffs")
@@
 summary(dat$playoffs)
-
-## playoffs variable
-#  Did Not make Playoffs or WC Play-in 
-#  Played in WC Play-In
-#  Eliminated Round 1
-#  Eliminated Round 2
-#  Lost WS
-#  Won WS
 ```
@@
 
-annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll, data = by_year) ## this should be an AR model
+# NOTE to self: check this again to see if i did it right
+
+annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll + lag(ghits_e1, 1), data = by_year) ## this should be an AR model, annual
+
+# i used the lag function , but examples I see have used dynlm package
 
@@
 Much more evened out for Dodgers and Dbacks.  Why not for Cubs?  Probably because
-of the later increasingtrend they have.
+of the later increasing trend they have.
 
@@
   dat  %>%
-  lm(ghits_e2 ~ phase + is_playoffs:playoffs, data = .) ## needs to be AR model instead
+  lm(ghits_e2 ~ phase + is_playoffs:playoffs + lag(ghits_e2, 1), data = .) ## needs to be AR model instead, monthly
 

### 2020-08-18 KB_modeling.Rmd
Added comments changes from AR term

@@
 
-## Added new dataset for playoff
+## Added new dataset for playoff: MLB_Postseason_Success.csv
 # broken up into 6 different factors
@@
 # NOTE to self: check this again to see if i did it right
-
-annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll + lag(ghits_e1, 1), data = by_year) ## this should be an AR model, annual
-
 # i used the lag function , but examples I see have used dynlm package
 
+annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll + lag(ghits_e1, 1), data = by_year) ## this should be an AR model
+# annual_means <- lm(ghits_e1 ~ odds + winpct + attendance + payroll, data = by_year)
+
 summary(annual_means)
-
 ```
@@
 Takeaway:  Payroll and winpct are a good predictor of annual ghits
+
+UPDATE: What changed after adding the AR term lagging one year?
+Looks like odds,  winpct, and the lag are all good predictors of annual ghits now.  Payroll is no longer significant.
 
@@
 
+UPDATE: What changed after adding AR term?
+Looks like heavier spikes on the under predictions? Note sure why the spikes are so sharp. Is that okay?  Looks more sporadic because the lag accounts for the yearly cycle.
+
 ---
@@
   lm(ghits_e2 ~ phase + is_playoffs:playoffs + lag(ghits_e2, 1), data = .) ## needs to be AR model instead, monthly
+  # lm(ghits_e2 ~ phase + is_playoffs:playoffs, data = .)
 
@@
 Takeaway:  More hits in bigger markets, as expected.
+
+UDPATE: Why is lost WS NA?  All variables are significant predictors of monthly ghits except winning WS.
 
@@
 
+UPDATE: What changed when I added the lag on the previous month?
+It seems like adding the lag term doesn't change this plot? Might be doing something wrong.
+
 ```{r}
@@
 ```
+
 ```{r}
@@
 
+UPDATE: AR did seem to help this underpredicting problem!
+The California team plot is much more sporadic.
+The histograms are centered closer to 0.
+And all the predictions seem to be closer than before!
+
 ---

### 2020-08-18 Paper_Outline.Rmd
outline of paper

---
title: "Modeling Bandwagoning and Hype in the MLB"
author: "Jake Esprabens and Kelly Bodwin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

[Abstract]
In this work, we model the Google search hits for MLB teams across a season as
a function of team success...

# Introduction

[General Stuff about bandwagoning]

The term "Bandwagon" is used in professional sports to refer to fans suddenly
showing interest in a team when that team achieves success - i.e., "jumping on 
the bandwagon" of the team.  ...

In this paper, we establish a hierarchical time series model that borrows across
data from all 30 MLB teams to predict internet search volume for a given team
in a given month.  We use these models to address the following questions:

1. Which MLB teams are more or less prone overall to attracting bandwagon fans?

2. When in recent MLB history were there periods of overhype or underhype for a
particular team?

## The Data

[where did we get it, etc]

[small table snippet of dataset]

Fig 1.

[Brief discussion: note how there are consistent phases of the season, 
but success strongly determines size of spike; e.g. Giants won in X years, 
Dodgers won in Y years]

### Google Trends Scaling

[The problem of how gtrends reports relative to max value]

Fig 2

Through brute force exploration, we determined that the global maximum was set by
searches for the Chicago Cubs in October 2016, during their historic World Series
victory run.  Thus, data for all teams was recorded relative to this maximum;
that is, a Google Hits measurement of 68 indicates that the team was searched
68% as often as "Chicago Cubs" in October 2016.

### Quantile Normalization

Because of the unusual relative scaling of the data, and the magnitude of the
spikes in interest during the World Series, this data is very skewed.

Fig 3: Histogram of all gtrends obserations.

The skew is too extreme to address with any common data transformations.  However,
we assert that the raw data values are not meaningful in and of themselves, as
they correspond to an arbitrary scale set by the most publicized month of baseball
during the time window of the data collection.  Therefore, we opted for a data 
transformation that retains the rank-ordering of the observations, but not the 
scaling.  Google Hits values were quantile normalized (reference) before model building.

### Season phases

A third challenge of Google Trends as a data source is that data is aggregated
monthly.  This allows us to track some nuance in annual patterns, but not at
a game-specific level.

Instead of predicting internet search volume by month, we include in our model
a variable we call *Phase*.  The season is divided into four phases:

* *Opening*: April
* *Season*: May - Sept
* *Playoffs*: October
* *Off-Season:* Nov - March

Fig 4

## Model-Building

## Measuring hype and bandwagoning

## Conclusions

## Future Work

### 2020-08-18 KB_modeling.Rmd
nothing

@@
 ```
+
 This is much better.

### 2020-08-19 KB_modeling.Rmd
changed lag from meeting

--- 2020-08-18 KB_modeling.Rmd
@@
   ggplot(aes(x = date, y = ghits_e2, color = team)) +
-  geom_line()
+  geom_line() +
 ```
@@
   dat  %>%
-  lm(ghits_e2 ~ phase + is_playoffs:playoffs + lag(ghits_e2, 1), data = .) ## needs to be AR model instead, monthly
+  lm(ghits_e2 ~ phase + is_playoffs:playoffs + lag(ghits_e2, 12), data = .) ## needs to be AR model instead, monthly
   # lm(ghits_e2 ~ phase + is_playoffs:playoffs, data = .)
@@
 dat %>%
-  filter(team %in% example_teams) %>%
+  filter(team == "Arizona Diamondbacks") %>%
+  # filter(team %in% example_teams) %>%
   ggplot(aes(x = date, y = ghits_e3, color = team)) +

### 2020-08-21 Paper_Outline.Rmd
Added some plots for the paper

@@
 knitr::opts_chunk$set(echo = FALSE)
+```
+
+```{r}
+library(tidyverse)
+library(gtrendsR)
 ```
@@
 
+```{r}
+fig1_dat <- read.csv("combined_data.csv") %>%
+  mutate(date = as.Date(date)) %>%
+  filter(team == "Los Angeles Dodgers" | team == "San Francisco Giants")
+
+fig1_dat %>%
+  ggplot(aes(x = as.Date(date), y = googlehits, group = team)) + 
+  geom_line(aes(color = team)) + 
+  theme_classic() +
+  labs(main="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
+  scale_color_manual(values=c("dodgerblue", "darkorange"))
+```
+
 [Brief discussion: note how there are consistent phases of the season, 
@@
 
-Fig 2
+Fig 2.
+
+```{r}
+fig2a_dat <- gtrends(keyword = c("%2Fm%2F04mjl", "%2Fm%2F0713r"),geo = "US", time = "all")$interest_over_time %>%
+  mutate(team = keyword,
+         googlehits = hits,
+         team = (gsub("%2Fm%2F04mjl","Los Angeles Dodgers", team)),
+         team = (gsub("%2Fm%2F0713r","San Francisco Giants", team))) %>%
+  select(date, googlehits, team)
+
+fig2a_dat %>%
+  ggplot(aes(x = as.Date(date), y = googlehits, group = team)) + 
+  geom_line(aes(color = team)) + 
+  theme_classic() +
+  labs(main="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
+  scale_color_manual(values=c("dodgerblue", "darkorange"))
+```
+
+```{r}
+fig2b_dat <- gtrends(keyword = c("%2Fm%2F04mjl", "%2Fm%2F01yjl"),geo = "US", time = "all")$interest_over_time %>%
+  mutate(team = keyword,
+         googlehits = hits,
+         team = (gsub("%2Fm%2F04mjl","Los Angeles Dodgers", team)),
+         team = (gsub("%2Fm%2F01yjl","Chicago Cubs", team))) %>%
+  select(date, googlehits, team)
+
+fig2b_dat %>%
+  ggplot(aes(x = as.Date(date), y = googlehits, group = team)) + 
+  geom_line(aes(color = team)) + 
+  theme_classic() +
+  labs(main="Google Trend of Dodgers and Cubs", x = "Date", y = "Google Hits") + 
+  scale_color_manual(values=c("firebrick2", "dodgerblue"))
+```
 

### 2020-08-22 Paper_Outline.Rmd
Added another graph

@@
   geom_line(aes(color = team)) + 
-  theme_classic() +
-  labs(main="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
+  theme_minimal() +
+  labs(title="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
   scale_color_manual(values=c("dodgerblue", "darkorange"))
@@
   geom_line(aes(color = team)) + 
-  theme_classic() +
-  labs(main="Google Trend of Dodgers and Cubs", x = "Date", y = "Google Hits") + 
+  theme_minimal() +
+  labs(title="Google Trend of Dodgers and Cubs", x = "Date", y = "Google Hits") + 
   scale_color_manual(values=c("firebrick2", "dodgerblue"))
@@
 Fig 3: Histogram of all gtrends obserations.
+
+```{r}
+fig3_dat <- read.csv("combined_data.csv")
+
+fig3_dat %>%
+  ggplot(aes(x = googlehits)) + 
+  geom_histogram(bins = 30,
+                 col = "dodgerblue4",
+                 fill = "dodgerblue",
+                 alpha = 0.7) +
+  theme_minimal() + 
+  theme(axis.title.y=element_blank()) + 
+  labs(main = "Google Trend Hits", x = "Google Hits")
+
+```
+
+```{r}
+fig3_dat %>%
+  ggplot(aes(x = log(googlehits))) + 
+  geom_histogram(bins = 10,
+                 col = "dodgerblue4",
+                 fill = "dodgerblue",
+                 alpha = 0.7) +
+  theme_minimal() + 
+  theme(axis.title.y=element_blank()) + 
+  labs(main = "Google Trend Hits", x = "Google Hits")
+```
 

### 2020-08-23 Paper_Outline.Rmd
Updated plots for paper outline

@@
   theme(axis.title.y=element_blank()) + 
-  labs(main = "Google Trend Hits", x = "Google Hits")
+  labs(title = "Google Trend Hits", x = "Google Hits")
 
@@
   theme(axis.title.y=element_blank()) + 
-  labs(main = "Google Trend Hits", x = "Google Hits")
+  labs(title = "Google Trend Hits", x = "Google Hits")
 ```
@@
 
+```{r}
+fig4_dat <- read.csv("plotting_data.csv") %>%
+  mutate(date = as.Date(date),
+         month = as.month(month))
+
+# giants - quantile normalized hits plot for last 5 years
+fig4_dat %>%
+  filter(team == "San Francisco Giants",
+         date <= as.Date("2020-07-01") & date >= as.Date("2015-07-01")) %>%
+  ggplot(aes(x = date, y = qn_ghits)) +
+  geom_line(col = "darkorange") +
+  theme_minimal() +
+  labs(title = "San Francisco Giants")
+
+# giants - monthly average hits over ALL years with phases
+fig4_dat %>%
+  filter(team == "San Francisco Giants") %>% 
+  group_by(month) %>%
+  summarise(monthly_avg = mean(googlehits)) %>%
+  ggplot(aes(x = month, y = monthly_avg)) +
+  geom_line(col = "darkorange") +
+  xlim(1,12) + 
+  geom_vline(xintercept = c(4,10), linetype = "dashed", color = "black") + 
+  geom_text(aes(x=2, y=7, label = "Off-Season")) +
+  geom_text(aes(x=4, y=7, label = "Opening Day\n"), angle = 90) + 
+  geom_text(aes(x=7, y=7, label = "Season")) + 
+  geom_text(aes(x=10, y=7, label = "Post-Season\n"), angle = 90) + 
+  geom_text(aes(x=11.5, y=7, label = "Off-Season"))
+  # geom_vline(xintercept = 10, linetype = "dashed", color = "black") + 
+  # geom_vline
+```
+
+```{r}
+# padres
+fig4_dat %>%
+  filter(team == "San Diego Padres") %>%
+  ggplot(aes(x = date, y = googlehits)) +
+  geom_line(col = "tan4") +
+  theme_minimal() +
+  labs(title = "San Diego Padres")
+```
+
 ## Model-Building

### 2020-08-24 Paper_Outline.Rmd
Updated graphs

--- 2020-08-21 Paper_Outline.Rmd
@@
   geom_line(aes(color = team)) + 
-  theme_classic() +
-  labs(main="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
-  scale_color_manual(values=c("dodgerblue", "darkorange"))
+  theme_minimal() +
+  labs(title="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
+  scale_color_manual(values=c("dodgerblue", "darkorange")) +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
 ```
 
-[Brief discussion: note how there are consistent phases of the season, 
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())[Brief discussion: note how there are consistent phases of the season, 
 but success strongly determines size of spike; e.g. Giants won in X years, 
@@
   geom_line(aes(color = team)) + 
-  theme_classic() +
-  labs(main="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
-  scale_color_manual(values=c("dodgerblue", "darkorange"))
+  theme_minimal() +
+  labs(title="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
+  scale_color_manual(values=c("dodgerblue", "darkorange")) +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
 ```
@@
   geom_line(aes(color = team)) + 
-  theme_classic() +
-  labs(main="Google Trend of Dodgers and Cubs", x = "Date", y = "Google Hits") + 
-  scale_color_manual(values=c("firebrick2", "dodgerblue"))
+  theme_minimal() +
+  labs(title="Google Trend of Dodgers and Cubs", x = "Date", y = "Google Hits") + 
+  scale_color_manual(values=c("firebrick2", "dodgerblue")) +
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+
 ```
@@
 Fig 3: Histogram of all gtrends obserations.
+
+```{r}
+fig3_dat <- read.csv("combined_data.csv")
+
+fig3_dat %>%
+  ggplot(aes(x = googlehits)) + 
+  geom_histogram(bins = 30,
+                 col = "dodgerblue4",
+                 fill = "dodgerblue",
+                 alpha = 0.7) +
+  theme_minimal() + 
+  theme(axis.title.y=element_blank()) + 
+  labs(title = "Google Trend Hits", x = "Google Hits")
+
+```
+
+```{r}
+fig3_dat %>%
+  ggplot(aes(x = log(googlehits))) + 
+  geom_histogram(bins = 10,
+                 col = "dodgerblue4",
+                 fill = "dodgerblue",
+                 alpha = 0.7) +
+  theme_minimal() + 
+  theme(axis.title.y=element_blank()) + 
+  labs(title = "Google Trend Hits", x = "Google Hits")
+```
 
@@
 
+```{r}
+fig4_dat <- read.csv("plotting_data.csv") %>%
+  mutate(date = as.Date(date))
+
+# giants - quantile normalized hits plot for last 5 years
+fig4_dat %>%
+  filter(team == "San Francisco Giants",
+         date <= as.Date("2020-07-01") & date >= as.Date("2015-07-01")) %>%
+  ggplot(aes(x = date, y = qn_ghits)) +
+  geom_line(col = "darkorange") +
+  theme_minimal() +
+  labs(title = "San Francisco Giants") + 
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+
+# giants - monthly average hits over ALL years with phases
+fig4_dat %>%
+  filter(team == "San Francisco Giants") %>% 
+  group_by(month) %>%
+  summarise(monthly_avg = mean(googlehits)) %>%
+  ggplot(aes(x = as.factor(month), y = monthly_avg, group = 1)) +
+  geom_point(col = "darkorange") +
+  geom_line(col = "darkorange") + 
+  theme_minimal() +
+  geom_vline(xintercept = c(4,10), linetype = "dashed", color = "black") + 
+  geom_text(aes(x=2, y=7, label = "Off-Season")) +
+  geom_text(aes(x=11.5, y=7, label = "Off-Season")) +
+  geom_text(aes(x=4, y=7, label = "Opening Day\n"), angle = 90) + 
+  geom_text(aes(x=7, y=7, label = "Season")) + 
+  geom_text(aes(x=10, y=7, label = "Post-Season\n"), angle = 90) + 
+  labs(title = "Monthly Average of Google Trend Hits for SF Giants (2004 - Present)") + 
+  scale_x_discrete(labels = month.abb[c(1:12)]) + 
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_blank())
+
+```
+
+```{r}
+# Mariners - quantile normalized hits plot for last 5 years
+fig4_dat %>%
+  filter(team == "Seattle Mariners",
+         date <= as.Date("2020-07-01") & date >= as.Date("2015-07-01")) %>%
+  ggplot(aes(x = date, y = qn_ghits)) +
+  geom_line(col = "cyan4") +
+  theme_minimal() +
+  labs(title = "Seattle Mariners") + 
+  theme(panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank())
+```
+
 ## Model-Building

### 2020-08-25 Paper_Outline.Rmd
Updated Plots and Abstract

--- 2020-08-22 Paper_Outline.Rmd
@@
 [Abstract]
-In this work, we model the Google search hits for MLB teams across a season as
-a function of team success...
+“Hopping on the bandwagon” is a familiar phrase used to describe a person that follows a trend or believes in an idea solely because it is becoming popular.
+The bandwagon effect is very common when describing fans in professional sports, including Major League Baseball.
+(Talk about Past Bandwagon Studies?) In this work, we model the Google search hits for MLB teams across a season as a function of team success.
+Using a hierarchical time series model, we are able to explore which MLB teams attract the most “bandwagon” fans and when MLB teams are overhyped or underhyped?
+We defined a “hype” moment as an unusual spike in Google searches at a specific time for a team.
+The creation of this model required gathering the Google search hits data from Google Trends for each MLB team and collecting other baseball-related data to be used as exogenous variables.
+(More Model Building) (Results and Conclusion)
 
@@
   geom_line(aes(color = team)) + 
-  theme_minimal() +
-  labs(title="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
-  scale_color_manual(values=c("dodgerblue", "darkorange"))
+  labs(title="Google Trend of LA Dodgers and SF Giants", x = "Date", y = "Google Hits") + 
+  scale_color_manual(values=c("dodgerblue", "darkorange")) +
+  theme_minimal()+
+  theme(legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))
 ```
@@
   labs(title="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
-  scale_color_manual(values=c("dodgerblue", "darkorange"))
+  scale_color_manual(values=c("dodgerblue", "darkorange")) +
+  theme(legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))
 ```
@@
   labs(title="Google Trend of Dodgers and Cubs", x = "Date", y = "Google Hits") + 
-  scale_color_manual(values=c("firebrick2", "dodgerblue"))
+  scale_color_manual(values=c("firebrick2", "dodgerblue")) +
+  theme(legend.position = "top",
+        legend.title = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))
+
 ```
@@
   ggplot(aes(x = googlehits)) + 
-  geom_histogram(bins = 30,
-                 col = "dodgerblue4",
-                 fill = "dodgerblue",
-                 alpha = 0.7) +
-  theme_minimal() + 
-  theme(axis.title.y=element_blank()) + 
-  labs(main = "Google Trend Hits", x = "Google Hits")
-
+  geom_histogram(bins = 20,
+                 col = "azure4",
+                 fill = "azure3") +
+  labs(title = "Google Trend Hits for All MLB Teams", x = "Google Hits") + 
+  scale_y_continuous(expand = c(0,0), limits = c(0, 4300)) +
+  theme(axis.title.y=element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA),
+        panel.background = element_rect(fill = "white"),
+        plot.background = element_rect(fill = "azure3")
+        ) 
 ```
@@
   geom_histogram(bins = 10,
-                 col = "dodgerblue4",
-                 fill = "dodgerblue",
-                 alpha = 0.7) +
-  theme_minimal() + 
-  theme(axis.title.y=element_blank()) + 
-  labs(main = "Google Trend Hits", x = "Google Hits")
+                 col = "azure4",
+                 fill = "azure3") +
+  labs(title = "Log Google Trend Hits for All MLB Teams", x = "Log Google Hits") + 
+  scale_y_continuous(expand = c(0,0), limits = c(0, 2200)) +
+  theme(axis.title.y=element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.major.y = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA),
+        panel.background = element_rect(fill = "white"),
+        plot.background = element_rect(fill = "azure3")
+        ) 
 ```
@@
 
+```{r}
+fig4_dat <- read.csv("plotting_data.csv") %>%
+  mutate(date = as.Date(date))
+
+# giants - quantile normalized hits plot for last 5 years
+fig4_dat %>%
+  filter(team == "San Francisco Giants",
+         date <= as.Date("2020-07-01") & date >= as.Date("2015-07-01")) %>%
+  ggplot(aes(x = date, y = qn_ghits)) +
+  geom_line(col = "darkorange") +
+  labs(title = "Quantile-Normalized Google Trend Hits for SF Giants (Past 5 Years)") + 
+  theme(plot.title = element_text(face = "bold"),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA),
+        panel.background = element_rect(fill = "gray90"),
+        plot.background = element_rect(fill = "darkorange"))
+
+# giants - monthly average hits over ALL years with phases
+fig4_dat %>%
+  filter(team == "San Francisco Giants") %>% 
+  group_by(month) %>%
+  summarise(monthly_avg = mean(googlehits)) %>%
+  ggplot(aes(x = as.factor(month), y = monthly_avg, group = 1)) +
+  geom_point(col = "darkorange") +
+  geom_line(col = "darkorange") + 
+  geom_vline(xintercept = c(4,10), linetype = "dashed", color = "black") + 
+  geom_text(aes(x=2, y=7, label = "Off-Season")) +
+  geom_text(aes(x=11.5, y=7, label = "Off-Season")) +
+  geom_text(aes(x=4, y=7, label = "Opening Day\n"), angle = 90) + 
+  geom_text(aes(x=7, y=7, label = "Season")) + 
+  geom_text(aes(x=10, y=7, label = "Post-Season\n"), angle = 90) + 
+  labs(title = "Monthly Average of Google Trend Hits for SF Giants (2004 - Present)") + 
+  scale_x_discrete(labels = month.abb[c(1:12)]) + 
+  theme(plot.title = element_text(face = "bold"),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA),
+        panel.background = element_rect(fill = "gray90"),
+        plot.background = element_rect(fill = "darkorange"))
+
+```
+
+```{r}
+# Mariners - quantile normalized hits plot for last 5 years
+fig4_dat %>%
+  filter(team == "Seattle Mariners",
+         date <= as.Date("2020-07-01") & date >= as.Date("2015-07-01")) %>%
+  ggplot(aes(x = date, y = qn_ghits)) +
+  geom_line(col = "cyan4") +
+  labs(title = "Quantile-Normalized Google Trend Hits for SEA Mariners (Past 5 Years)") + 
+  theme(plot.title = element_text(face = "bold"),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA),
+        panel.background = element_rect(fill = "gray90"),
+        plot.background = element_rect(fill = "cyan4"))
+
+# Mariners - monthly average hits over ALL years with phases
+fig4_dat %>%
+  filter(team == "Seattle Mariners") %>% 
+  group_by(month) %>%
+  summarise(monthly_avg = mean(googlehits)) %>%
+  ggplot(aes(x = as.factor(month), y = monthly_avg, group = 1)) +
+  geom_point(col = "cyan4") +
+  geom_line(col = "cyan4") + 
+  geom_vline(xintercept = c(4,10), linetype = "dashed", color = "black") + 
+  geom_text(aes(x=2, y=2, label = "Off-Season")) +
+  geom_text(aes(x=11.5, y=2, label = "Off-Season")) +
+  geom_text(aes(x=4, y=2, label = "Opening Day\n"), angle = 90) + 
+  geom_text(aes(x=7, y=2, label = "Season")) + 
+  geom_text(aes(x=10, y=2, label = "Post-Season\n"), angle = 90) + 
+  labs(title = "Monthly Average of Google Trend Hits for SEA Mariners (2004 - Present)") + 
+  scale_x_discrete(labels = month.abb[c(1:12)]) + 
+  theme(plot.title = element_text(face = "bold"),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA),
+        panel.background = element_rect(fill = "gray90"),
+        plot.background = element_rect(fill = "cyan4"))
+
+```
+
+```{r}
+# ALL TEAMS - quantile normalized hits plot for last 5 years
+fig4_dat %>%
+  filter(date <= as.Date("2020-07-01") & date >= as.Date("2015-07-01")) %>%
+  group_by(date) %>%
+  summarise(avg_qn_ghits = mean(qn_ghits)) %>%
+  ggplot(aes(x = date, y = avg_qn_ghits)) +
+  geom_line(col = "azure4") +
+  labs(title = "Average Quantile-Normalized Google Trend Hits for MLB Teams (Past 5 Years)") + 
+  theme(plot.title = element_text(face = "bold", size = 13),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.grid.minor.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA),
+        panel.background = element_rect(fill = "gray90"),
+        plot.background = element_rect(fill = "azure3"))
+
+# ALL TEAMS - monthly average hits over ALL years with phases
+fig4_dat %>%
+  group_by(month) %>%
+  summarise(monthly_avg = mean(googlehits)) %>%
+  ggplot(aes(x = as.factor(month), y = monthly_avg, group = 1)) +
+  geom_point(col = "azure4") +
+  geom_line(col = "azure4") + 
+  geom_vline(xintercept = c(4,10), linetype = "dashed", color = "black") + 
+  geom_text(aes(x=2, y=3.5, label = "Off-Season")) +
+  geom_text(aes(x=11.5, y=3.5, label = "Off-Season")) +
+  geom_text(aes(x=4, y=3.5, label = "Opening Day\n"), angle = 90) + 
+  geom_text(aes(x=7, y=3.5, label = "Season")) + 
+  geom_text(aes(x=10, y=3.5, label = "Post-Season\n"), angle = 90) + 
+  labs(title = "Monthly Average of Google Trend Hits for MLB Teams (2004 - Present)") + 
+  scale_x_discrete(labels = month.abb[c(1:12)]) + 
+  theme(plot.title = element_text(face = "bold"),
+        axis.title.x = element_blank(),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA),
+        panel.background = element_rect(fill = "gray90"),
+        plot.background = element_rect(fill = "azure3"))
+
+```
+
 ## Model-Building

### 2020-08-25 KB_modeling.Rmd
Updated plots for paper outline

@@
 
+write.csv(dat, "C:/Users/espra/OneDrive/Documents/CP/Research - Bandwagon/MLB-fan-support/plotting_data.csv")
 ```

### 2020-08-25 Paper_Outline.Rmd
Fixed a little formatting
Fixed some typos in outline

@@
 [Abstract]
-“Hopping on the bandwagon” is a familiar phrase used to describe a person that follows a trend or believes in an idea solely because it is becoming popular.
-The bandwagon effect is very common when describing fans in professional sports, including Major League Baseball.
-(Talk about Past Bandwagon Studies?) In this work, we model the Google search hits for MLB teams across a season as a function of team success.
-Using a hierarchical time series model, we are able to explore which MLB teams attract the most “bandwagon” fans and when MLB teams are overhyped or underhyped?
-We defined a “hype” moment as an unusual spike in Google searches at a specific time for a team.
-The creation of this model required gathering the Google search hits data from Google Trends for each MLB team and collecting other baseball-related data to be used as exogenous variables.
-(More Model Building) (Results and Conclusion)
+“Hopping on the bandwagon” is a familiar phrase used to describe a person that 
+follows a trend, believes in an idea, or supports a team solely because it is 
+becoming popular. The bandwagon effect is very common when describing fans in 
+professional sports, including Major League Baseball. (Talk about Past Bandwagon 
+Studies?) In this work, we model the Google search hits for MLB teams across a 
+season as a function of team success.  Using a hierarchical time series model, we 
+are able to explorewhich MLB teams attract the most “bandwagon” fans and when MLB 
+teams are overhypedor underhyped. We defined a “hype” moment as an unusual spike 
+in Google searches at a specific time for a team.  The creation of this model 
+required gathering the Google search hits data from Google Trends for each MLB 
+team and collecting other baseball-related data to be used as exogenous 
+variables. (...More Model Building) (...Results and Conclusion)
 

### 2020-08-25 KB_model_2.Rmd
model updates

---
title: "Untitled"
author: "Kelly Bodwin"
date: "8/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs}
library(tidyverse)
library(tidymodels)
```

```{r read data}
quant_norm <- function(vec){

  quants <- rank(vec)/(length(vec) + 1)
  normies <- purrr::map_dbl(quants, qnorm)

  return(normies)

}

dat_orig <- read_csv(here::here("MLB_combined.csv")) 
```

```{r data prep}
mlb_hype <- read_csv("mlb_hype.csv")

mlb_hype <- mlb_hype %>%
  mutate(
    Month = paste0(Month, "-01"),
    date = ymd(Month)
  )

dat <- dat_orig %>%
  left_join(mlb_hype) %>%
  mutate(
    googlehits_orig = googlehits,
    googlehits = quant_norm(googlehits),
    mlb_searches_orig = mlb_searches,
    mlb_searches = quant_norm(mlb_searches),
    playoffs = factor(playoffs, 
                      levels = c("No Playoffs", "Lost WC", "Lost Quarter", "Lost Semi", "Lost WS", "Won WS"))
  )

# 
# dat_split <- initial_split(dat)
# 
# train_dat <- training(dat_split)
# test_dat <- testing(dat_split)
```

## Notes

Why might a team have higher google hits in a given month?

* *Market Size*:  More fans = more people googling.

* *Leaguewide Interest*:  More people follow baseball during the playoffs.

* *Prior Performance:*  A team that wins the WS will be followed more closely the next year.

* *Expectations*: A team that is considered likely to content is in the news more.

* *Recent Performance:* A team on a hot streak will be in the news.

* *Special Events:* A trade, an off-field scandal, etc.

Predict each month based on these things?

```{r}
dat %>%
  mutate(
    mlb_searches = quant_norm(mlb_searches)
  ) %>%
  filter(team %in% c("San Francisco Giants", "Chicago Cubs")) %>%
  ggplot(aes(x = date, y = googlehits, color = team)) +
  geom_line() +
  geom_line(aes(x = date, y = mlb_searches), color = "black")
```

```{r}
dat_overall <- dat %>%
  group_by(team) %>%
  summarize(
    ghits_overall = mean(googlehits),
    propTVMarketSize = propTVMarketSize[1]
  ) %>%
  ungroup()

dat_annual <- dat %>%
  group_by(team, year) %>%
  summarize(
    ghits_annual = mean(googlehits),
    mlb_annual = mean(mlb_searches),
    propTVMarketSize = propTVMarketSize[1],
    playoffs = playoffs[1],
    odds = odds[1],
    winpct = winpct[1],
    attendance = attendance[1],
    payroll = payroll[1]
  ) %>%
  ungroup()

dat_annual <- dat_overall %>%
  select(team, ghits_overall) %>%
  right_join(dat_annual)

dat <- dat_annual %>%
  select(team, year, ghits_annual, mlb_annual, ghits_overall) %>%
  right_join(dat)
```

```{r overall model}
fit_overall <- lm(ghits_overall ~ propTVMarketSize,
                      data = dat_overall)

dat_annual <- dat_annual %>%
  add_predictions(
    fit_overall,
    var = "pred_overall"
  ) %>%
  mutate(
    resid_annual = ghits_annual - pred_overall
  )
```

```{r annual model}
fit_annual <- lm(resid_annual ~ mlb_annual + 
                   lag(ghits_annual) + 
                   lag(playoffs) + lag(winpct) + lag(attendance) +
                   odds + payroll,
                      data = dat_annual)

dat <- dat %>%
  add_predictions(
    fit_overall,
    var = "pred_overall"
  ) %>%
  add_predictions(
    fit_annual,
    var = "pred_resid_annual"
  ) %>%
  mutate(
    pred_annual = pred_overall + pred_resid_annual,
    resid = googlehits - pred_annual
  )
```

```{r monthly model}
fit_monthly <- lm(resid ~ mlb_searches + lag(googlehits) + 
                        phase*(winpct + attendance) + is_playoffs:playoffs,
                      data = dat)

dat <- dat %>%
  add_predictions(
    fit_monthly,
    var = "pred_resid"
  ) %>%
  mutate(
    pred = pred_overall + pred_resid_annual + pred_resid,
    resid_final = googlehits - pred
  )
```

```{r final resids}
dat %>%
  filter(team %in% c("San Francisco Giants", "Los Angeles Dodgers", "San Diego Padres")) %>%
  ggplot(aes(x = date, y = resid_final, color = team)) +
  geom_line()

dat %>%
  filter(team %in% c("San Francisco Giants")) %>%
  ggplot() +
  geom_line(aes(x = date, y = googlehits), color = "black") +
  geom_line(aes(x = date, y = pred), color = "red")

dat %>%
  filter(team %in% c("Chicago Cubs")) %>%
  ggplot() +
  geom_line(aes(x = date, y = googlehits), color = "black") +
  geom_line(aes(x = date, y = pred), color = "red")
```

$$\text{\{Google Hits\}}_i \sim \alpha_i \, \text{\{TV Market Size\}_i} + E_1 + \epsilon_1$$

where $\epsilon \sim N(0, \sigma^2)$

$$E_{1ij} \sim \gamma_{ij} ... + E_2$$

## Team-specific

```{r}
lm(resid_final ~ team:phase, data = dat)
```

## Extra

```{r}
dat %>%
  filter(phase == "Playoffs") %>%
  lm(googlehits ~ propTVMarketSize+ playoffs + winpct, data = .) %>% 
  tidy() %>%
  ggplot(aes(x = fct_inorder(term), y = estimate)) +
  geom_col()

dat %>%
  filter(phase == "Season") %>%
  lm(googlehits ~ propTVMarketSize + attendance + odds + winpct, data = .) %>% 
  tidy() %>%
  ggplot(aes(x = fct_inorder(term), y = estimate)) +
  geom_col()
```

```{r}
by_phase <- dat %>%
  mutate(
    year = year(date)
  ) %>%
  drop_na() %>%
  group_by(team, year) %>%
  mutate(googlehits = scale(googlehits)) %>%
  group_by(team, year, phase) %>%
  summarize(hits = mean(googlehits)) %>%
  pivot_wider(names_from = phase, 
              values_from = hits) %>%
  ungroup()

plot(by_phase$Playoffs, by_phase$Season)

by_phase %>%
  select(Opening, Playoffs, Season, `Off-Season`) %>%
  drop_na() %>%
  as.matrix() %>%
  cor()
```

### 2020-08-25 KB_modeling.Rmd
model updates

--- 2020-08-18 KB_modeling.Rmd
@@
 
-playoff_success <- read.csv("MLB_Postseason_Success.csv", stringsAsFactors = FALSE) %>%
+playoff_success <- read_csv("MLB_Postseason_Success.csv") %>%
   mutate_all(na_if,"") %>%
@@
 #combining new palyoff dataset with dat
-dat <- left_join(dat, playoff_success, by = c("team", "year")) %>%
-  select(-c(DivWin, WCWin, LgWin, WSWin)) # deselecting unnecessary variables
+dat <- dat %>% 
+  left_join(playoff_success, by = c("team", "year"))
 
 dat <- dat %>%
-  # mutate_at(vars(DivWin, WCWin, LgWin, WSWin), funs(as.logical)) %>%
+  select(-c(DivWin, WCWin, LgWin, WSWin)) %>%
   mutate(
@@
     phase = case_when(
-      month < 04 | month > 10 ~ "Summer",
+      month < 04 | month > 10 ~ "Off-Season",
       month == 04 ~ "Opening",
@@
 summary(dat$playoffs)
+
+dat %>%
+  write_csv("MLB_combined.csv")
 ```

### 2020-08-27 Paper_Outline.Rmd
Updated graphs a bit

@@
   theme_minimal()+
-  theme(legend.position = "top",
-        legend.title = element_blank(),
+  annotate(geom="text",x=as.Date("2012-10-01"), y=33,label="SF Giants", col = "darkorange", fontface = "bold", size = 3.5) + 
+  annotate(geom="text",x=as.Date("2017-01-01"), y=52,label="LA Dodgers", col = "dodgerblue", fontface = "bold", size = 3.5) + 
+  theme(legend.position = "none",
         panel.grid.major.x = element_blank(),
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
+        # panel.grid.minor.y = element_blank(),
+        panel.border = element_rect(colour = "gray85", fill = NA))
 ```
@@
   theme_minimal() +
-  labs(title="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
+  labs(title="Google Trend of LA Dodgers and SF Giants", x = "Date", y = "Google Hits") + 
   scale_color_manual(values=c("dodgerblue", "darkorange")) +
-  theme(legend.position = "top",
-        legend.title = element_blank(),
+  annotate(geom="text",x=as.Date("2012-10-01"), y=45,label="SF Giants", col = "darkorange", fontface = "bold", size = 3.5) + 
+  annotate(geom="text",x=as.Date("2017-01-01"), y=80,label="LA Dodgers", col = "dodgerblue", fontface = "bold", size = 3.5) + 
+  theme(legend.position = "none",
         panel.grid.major.x = element_blank(),
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
+        panel.border = element_rect(colour = "gray85", fill = NA))
 ```
@@
   theme_minimal() +
-  labs(title="Google Trend of Dodgers and Cubs", x = "Date", y = "Google Hits") + 
+  labs(title="Google Trend of LA Dodgers and CHI Cubs", x = "Date", y = "Google Hits") + 
   scale_color_manual(values=c("firebrick2", "dodgerblue")) +
-  theme(legend.position = "top",
-        legend.title = element_blank(),
+  annotate(geom="text",x=as.Date("2015-06-01"), y=50,label="CHI Cubs", col = "firebrick2", fontface = "bold", size = 3.5) + 
+  annotate(geom="text",x=as.Date("2019-01-01"), y=70,label="LA Dodgers", col = "dodgerblue", fontface = "bold", size = 3.5) + 
+  theme(legend.position = "none",
         panel.grid.major.x = element_blank(),
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
+        # panel.grid.major.y = element_blank(),
+        panel.grid.minor.y = element_blank(),
+        panel.border = element_rect(colour = "gray85", fill = NA))
 
@@
         panel.grid.minor.y = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA),
+        panel.border = element_rect(colour = "gray85", fill = NA),
         panel.background = element_rect(fill = "white"),
-        plot.background = element_rect(fill = "azure3")
+        # plot.background = element_rect(fill = "azure3")
         ) 

### 2020-08-28 Paper_Outline.Rmd
Updated plots more and added to paper

--- 2020-08-25 Paper_Outline.Rmd
@@
 
+### Google hits
+In order to measure the bandwagon level of MLB fans, we gathered data from google trends to depict the popularity of each team.
+This data was gathered using the [**gtrendsR**](https://cran.r-project.org/web/packages/gtrendsR/gtrendsR.pdf) package, which automatically pulls the data from the [Google Trends](https://trends.google.com/trends/?geo=US) API.
+The units of the Google Trend data are dependent on the maximum of all the Google searches of interest.
+Therefore, we went through each team and determined that the Chicago Cubs had the maximum google trend hits; this lead us to gather data for the rest of the MLB teams relative to the Cubs.
+We were also able to collect the Google search hits of the term "MLB" using Google Trends.
+
+### Sports Betting Odds
+The preseason odds dataset was gathered using data from [Sports Odds History](https://www.sportsoddshistory.com/mlb-odds/).
+We compiled this data by viewing each teams' odds of winning the World Series from 2004 to the present.
+Ideally, we would have like to use monthly odds of winning the World Series for each MLB team; however, this data was not available for every desired year.
+
+### Attendance and Payroll
+We decided that attendence and payroll of each team by season may be valuable exogenous variables that could help determine when fans would become bandwagons.
+We gathered this data from [Baseball Reference](https://www.baseball-reference.com/leagues/MLB/2019-misc.shtml) by copying the data from each the years 2004 to 2019.
+
+### Team Performance
+Other exogenous variables we decided would be a valuable predictor in this model was win percentage and whether each team made the playoffs.
+We were able to download this data directly from [Sean Lahman's Baseball Database](http://www.seanlahman.com/baseball-archive/statistics/).
+This dataset was also filled with other variables such as attendance along with indicators of whether each team won their respective divisions, wild card games, leagues, and the world series.
+
+### TV Market Size
+The TV Market Size of each team is was also important to create an accurate estimation of the number of true fans of each.
+We were able to gather this data using the Nielsen Market Size from [Sports Media Watch](https://www.sportsmediawatch.com/nba-market-size-nfl-mlb-nhl-nielsen-ratings/).
+
+### Postseason Performance
+Postseason performance is likely to be one of the most, if not the most, valuable exogenous predictors a teams Google search hits.
+We manually gathered this data using the respective [Wikipedia](https://en.wikipedia.org/wiki/2019_Major_League_Baseball_season#Postseason) page for each MLB postseason.
+
 [small table snippet of dataset]
+```{r, echo = FALSE}
+head(read.csv("MLB_combined.csv"), 10)
+```
 
@@
   geom_line(aes(color = team)) + 
-  labs(title="Google Trend of LA Dodgers and SF Giants", x = "Date", y = "Google Hits") + 
+  labs(title="Google Trend of LA Dodgers and SF Giants (2004-2020)", x = "Date", y = "Google Hits") + 
   scale_color_manual(values=c("dodgerblue", "darkorange")) +
   theme_minimal()+
-  theme(legend.position = "top",
-        legend.title = element_blank(),
+  annotate(geom="text",x=as.Date("2012-10-01"), y=33,label="SF Giants", col = "darkorange", fontface = "bold", size = 3.5) + 
+  annotate(geom="text",x=as.Date("2017-01-01"), y=52,label="LA Dodgers", col = "dodgerblue", fontface = "bold", size = 3.5) + 
+  theme(legend.position = "none",
         panel.grid.major.x = element_blank(),
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(angle = 0, vjust = 0.5),
+        panel.border = element_rect(colour = "gray85", fill = NA))
 ```
@@
   theme_minimal() +
-  labs(title="Google Trend of Dodgers and Giants", x = "Date", y = "Google Hits") + 
+  labs(title="Google Trend of LA Dodgers and SF Giants (2004-2020)", x = "Date", y = "Google Hits") + 
   scale_color_manual(values=c("dodgerblue", "darkorange")) +
-  theme(legend.position = "top",
-        legend.title = element_blank(),
+  annotate(geom="text",x=as.Date("2012-10-01"), y=45,label="SF Giants", col = "darkorange", fontface = "bold", size = 3.5) + 
+  annotate(geom="text",x=as.Date("2017-01-01"), y=80,label="LA Dodgers", col = "dodgerblue", fontface = "bold", size = 3.5) + 
+  theme(legend.position = "none",
         panel.grid.major.x = element_blank(),
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
+        axis.title.x = element_blank(),
+        axis.title.y = element_text(angle = 0, vjust = 0.5),
+        panel.border = element_rect(colour = "gray85", fill = NA))
 ```
@@
   theme_minimal() +
-  labs(title="Google Trend of Dodgers and Cubs", x = "Date", y = "Google Hits") + 
+  labs(title="Google Trend of LA Dodgers and CHI Cubs (2004-2020)", x = "Date", y = "Google Hits") + 
   scale_color_manual(values=c("firebrick2", "dodgerblue")) +
-  theme(legend.position = "top",
-        legend.title = element_blank(),
+  annotate(geom="text",x=as.Date("2015-06-01"), y=50,label="CHI Cubs", col = "firebrick2", fontface = "bold", size = 3.5) + 
+  annotate(geom="text",x=as.Date("2019-01-01"), y=70,label="LA Dodgers", col = "dodgerblue", fontface = "bold", size = 3.5) + 
+  theme(legend.position = "none",
         panel.grid.major.x = element_blank(),
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
+        axis.title.y = element_text(angle = 0, vjust = 0.5),
+        axis.title.x = element_blank(),
+        panel.border = element_rect(colour = "gray85", fill = NA))
 
@@
   theme(axis.title.y=element_blank(),
-        panel.grid.major.x = element_blank(),
-        panel.grid.major.y = element_blank(),
-        panel.grid.minor.x = element_blank(),
-        panel.grid.minor.y = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA),
-        panel.background = element_rect(fill = "white"),
-        plot.background = element_rect(fill = "azure3")
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.background = element_rect(fill = "white")
         ) 
@@
   labs(title = "Quantile-Normalized Google Trend Hits for SF Giants (Past 5 Years)") + 
-  theme(plot.title = element_text(face = "bold"),
+  theme(plot.title = element_text(face = "bold", size = 13),
         axis.title.x = element_blank(),
@@
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA),
-        panel.background = element_rect(fill = "gray90"),
+        panel.grid.major.y = element_line(colour = "gray85"),
+        panel.grid.minor.y = element_line(colour = "gray85"),
+        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.background = element_rect(fill = "white"),
         plot.background = element_rect(fill = "darkorange"))
@@
   geom_vline(xintercept = c(4,10), linetype = "dashed", color = "black") + 
-  geom_text(aes(x=2, y=7, label = "Off-Season")) +
-  geom_text(aes(x=11.5, y=7, label = "Off-Season")) +
-  geom_text(aes(x=4, y=7, label = "Opening Day\n"), angle = 90) + 
-  geom_text(aes(x=7, y=7, label = "Season")) + 
-  geom_text(aes(x=10, y=7, label = "Post-Season\n"), angle = 90) + 
-  labs(title = "Monthly Average of Google Trend Hits for SF Giants (2004 - Present)") + 
+  annotate(geom = "text", x=2, y=6.5, label = "Off-Season") +
+  annotate(geom = "text", x=4, y=6.5, label = "Opening Day\n", angle = 90) + 
+  annotate(geom = "text", x=7, y=6.5, label = "Season") +
+  annotate(geom = "text", x=10, y=6.5, label = "Post-Season\n", angle = 90) +
+  annotate(geom = "text", x=11.5, y=6.5, label = "Off-Season") +
+  labs(title = "Monthly Average of Google Trend Hits for SF Giants (2004-2020)") + 
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
@@
         panel.grid.major.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA),
-        panel.background = element_rect(fill = "gray90"),
+        panel.grid.major.y = element_line(colour = "gray85"),
+        panel.grid.minor.y = element_line(colour = "gray85"),
+        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.background = element_rect(fill = "white"),
         plot.background = element_rect(fill = "darkorange"))
@@
   labs(title = "Quantile-Normalized Google Trend Hits for SEA Mariners (Past 5 Years)") + 
-  theme(plot.title = element_text(face = "bold"),
+  theme(plot.title = element_text(face = "bold", size = 13),
         axis.title.x = element_blank(),
@@
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA),
-        panel.background = element_rect(fill = "gray90"),
+        panel.grid.major.y = element_line(colour = "gray85"),
+        panel.grid.minor.y = element_line(colour = "gray85"),
+        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.background = element_rect(fill = "white"),
         plot.background = element_rect(fill = "cyan4"))
@@
   geom_vline(xintercept = c(4,10), linetype = "dashed", color = "black") + 
-  geom_text(aes(x=2, y=2, label = "Off-Season")) +
-  geom_text(aes(x=11.5, y=2, label = "Off-Season")) +
-  geom_text(aes(x=4, y=2, label = "Opening Day\n"), angle = 90) + 
-  geom_text(aes(x=7, y=2, label = "Season")) + 
-  geom_text(aes(x=10, y=2, label = "Post-Season\n"), angle = 90) + 
-  labs(title = "Monthly Average of Google Trend Hits for SEA Mariners (2004 - Present)") + 
+  annotate(geom = "text", x=2, y=1.75, label = "Off-Season") +
+  annotate(geom = "text", x=4, y=1.75, label = "Opening Day\n", angle = 90) + 
+  annotate(geom = "text", x=7, y=1.75, label = "Season") +
+  annotate(geom = "text", x=10, y=1.75, label = "Post-Season\n", angle = 90) +
+  annotate(geom = "text", x=11.5, y=1.75, label = "Off-Season") +
+  labs(title = "Monthly Average of Google Trend Hits for SEA Mariners (2004-2020)") + 
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
@@
         panel.grid.major.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA),
-        panel.background = element_rect(fill = "gray90"),
+        panel.grid.major.y = element_line(colour = "gray85"),
+        panel.grid.minor.y = element_line(colour = "gray85"),
+        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.background = element_rect(fill = "white"),
         plot.background = element_rect(fill = "cyan4"))
@@
         panel.grid.minor.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA),
-        panel.background = element_rect(fill = "gray90"),
+        panel.grid.major.y = element_line(colour = "gray85"),
+        panel.grid.minor.y = element_line(colour = "gray85"),
+        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.background = element_rect(fill = "white"),
         plot.background = element_rect(fill = "azure3"))
@@
   geom_vline(xintercept = c(4,10), linetype = "dashed", color = "black") + 
-  geom_text(aes(x=2, y=3.5, label = "Off-Season")) +
-  geom_text(aes(x=11.5, y=3.5, label = "Off-Season")) +
-  geom_text(aes(x=4, y=3.5, label = "Opening Day\n"), angle = 90) + 
-  geom_text(aes(x=7, y=3.5, label = "Season")) + 
-  geom_text(aes(x=10, y=3.5, label = "Post-Season\n"), angle = 90) + 
-  labs(title = "Monthly Average of Google Trend Hits for MLB Teams (2004 - Present)") + 
+  annotate(geom = "text", x=2, y=2.5, label = "Off-Season") +
+  annotate(geom = "text", x=4, y=2.5, label = "Opening Day\n", angle = 90) + 
+  annotate(geom = "text", x=7, y=2.5, label = "Season") +
+  annotate(geom = "text", x=10, y=2.5, label = "Post-Season\n", angle = 90) +
+  annotate(geom = "text", x=11.5, y=2.5, label = "Off-Season") +
+  labs(title = "Monthly Average of Google Trend Hits for MLB Teams (2004-2020)") + 
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
@@
         panel.grid.major.x = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA),
-        panel.background = element_rect(fill = "gray90"),
+        panel.grid.major.y = element_line(colour = "gray85"),
+        panel.grid.minor.y = element_line(colour = "gray85"),
+        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.background = element_rect(fill = "white"),
         plot.background = element_rect(fill = "azure3"))

### 2020-08-29 Paper_Outline.Rmd
Updated paper and plots

@@
 ### Google hits
-In order to measure the bandwagon level of MLB fans, we gathered data from google trends to depict the popularity of each team.
+In order to measure the "bandwagon" level of MLB fans, we gathered data from Google Trends to depict the popularity of each team.
 This data was gathered using the [**gtrendsR**](https://cran.r-project.org/web/packages/gtrendsR/gtrendsR.pdf) package, which automatically pulls the data from the [Google Trends](https://trends.google.com/trends/?geo=US) API.
@@
 We were also able to collect the Google search hits of the term "MLB" using Google Trends.
+Moreover, it is important to note that Google Trends has only been keeping track of Google Searches since the beginning of 2004.
 
@@
 ### TV Market Size
-The TV Market Size of each team is was also important to create an accurate estimation of the number of true fans of each.
+The TV Market Size of each team was also important to create an accurate estimation of the number of true fans of each.
 We were able to gather this data using the Nielsen Market Size from [Sports Media Watch](https://www.sportsmediawatch.com/nba-market-size-nfl-mlb-nhl-nielsen-ratings/).
@@
 
+Here are two Google Trend lines of the San Francisco Giants and Los Angeles Dodgers from 2004 to July of 2020.
+Generally, it is evident that there is seasonality.
+There are often more Google search hits at the beginning of the season (April) and during the post-season (October); and expectedly, this number drastically drops during the off-season, which is from November to March.
+However, there are several out-of-the-ordinary substantial spikes in the last decade for both teams.
+These spikes are most strongly influenced by their playoff success.
+It is important to note that the San Francisco Giants won the World Series in 2010, 2012, and 2014 while the Dodgers made the World Series in 2017 and 2018, but failed to win.
+This clearly aligns with these sizable spikes.
+
 ### Google Trends Scaling
@@
 [The problem of how gtrends reports relative to max value]
+
+The Google Trends data is very useful; however, it can sometimes be hard to interpret because of the units.
+Google normalizes the data by dividing each data point by the total searches of the region and time range represented.
+In our case, the region is simply the United States and time varies from 2004 to 2020.
+These numbers are then put on scale of 0 to 100 relative to the rest of the correponding search terms.
+
+To illustrate how this process works, there are two plots below: Figure 2a compares just the Los Angeles Dodgers and the San Francisco Giants while Figure 2b only compares the Los Angeles Dodgers and the Chicago Cubs.
+In Figure 2a, it is evident that the Dodgers held the most Google hits in October of 2018 hence the spike reaching 100 Google hits.
+Every other point is scaled relative to this maximum.
+As we can see in Figure 2b, the scale will change for specific teams depending on the maximum.
+In this plot, the Cubs now hold the new maximum of 100 in October 2016 because this was the first time they had won the World Series since 1908.
+Therefore, they became the most googled team since 2004 resulting in the Dodgers no longer holding the maximum of 100 in this trend.
 
@@
 
-The skew is too extreme to address with any common data transformations.  However,
-we assert that the raw data values are not meaningful in and of themselves, as
-they correspond to an arbitrary scale set by the most publicized month of baseball
-during the time window of the data collection.  Therefore, we opted for a data 
-transformation that retains the rank-ordering of the observations, but not the 
+As seen in Figure 3a and 3b, the skew is too extreme to address with any common data 
+transformations.  However, we assert that the raw data values are not meaningful in 
+and of themselves, as they correspond to an arbitrary scale set by the most publicized 
+month of baseball during the time window of the data collection.  Therefore, we opted 
+for a data transformation that retains the rank-ordering of the observations, but not the 
 scaling.  Google Hits values were quantile normalized (reference) before model building.
@@
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
-  theme(plot.title = element_text(face = "bold"),
+  scale_y_discrete(expand = c(0,0), limits = factor(0:9)) +
+  theme(
+      # plot.title = element_text(face = "bold"),
         axis.title.x = element_blank(),
@@
         panel.grid.major.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "gray85"),
-        panel.grid.minor.y = element_line(colour = "gray85"),
-        panel.border = element_rect(colour = "gray85", fill = NA),
-        panel.background = element_rect(fill = "white"),
-        plot.background = element_rect(fill = "darkorange"))
+        panel.grid.major.y = element_line(colour = "gray90"),
+        panel.grid.minor.y = element_blank(),
+        panel.border = element_rect(colour = "gray75", fill = NA),
+        panel.background = element_rect(fill = "white")
+        # plot.background = element_rect(fill = "darkorange")
+  )
 
@@
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
-  theme(plot.title = element_text(face = "bold"),
+  scale_y_continuous(expand = c(0,0), limits = c(0, 3.5)) +
+  theme(
+        # plot.title = element_text(face = "bold"),
         axis.title.x = element_blank(),
@@
         panel.grid.major.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "gray85"),
-        panel.grid.minor.y = element_line(colour = "gray85"),
-        panel.border = element_rect(colour = "gray85", fill = NA),
-        panel.background = element_rect(fill = "white"),
-        plot.background = element_rect(fill = "cyan4"))
+        panel.grid.major.y = element_line(colour = "gray90"),
+        panel.grid.minor.y = element_line(colour = "gray90"),
+        panel.border = element_rect(colour = "gray80", fill = NA),
+        panel.background = element_rect(fill = "white")
+        # plot.background = element_rect(fill = "cyan4")
+        )
 
@@
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
+  scale_y_continuous(expand = c(0,0), limits = c(0, 5)) +
   theme(plot.title = element_text(face = "bold"),
@@
         panel.grid.major.y = element_line(colour = "gray85"),
-        panel.grid.minor.y = element_line(colour = "gray85"),
+        panel.grid.minor.y = element_blank(),
         panel.border = element_rect(colour = "gray85", fill = NA),
-        panel.background = element_rect(fill = "white"),
-        plot.background = element_rect(fill = "azure3"))
+        panel.background = element_rect(fill = "white")
+        # plot.background = element_rect(fill = "azure3")
+        )
 

### 2020-08-31 Paper_Outline.Rmd
more model details

@@
 
-## Measuring hype and bandwagoning
+(motivation for hierarchical model)
+
+(for sections below, provide discussion of model in words, and summary of model results)
+
+### Level 1: Overall predicted mean
+
+$$\{\text{Mean google hits overall} \} \sim \{\text{Mean google hits for all teams} \} + \alpha \, \text{\{TV Market Size\}} + \epsilon_{i}$$
+
+### Level 2: Annual predicted mean
+
+$$\{\text{Mean google hits this year} \} \sim 
+\{\text{Mean google hits overall} \} + 
+\gamma_0 \{\text{MLB google hits this year}\} + 
+\gamma_{1} \{\text{Mean google hits last year}\} +
+\gamma_{2} \{\text{Playoffs last year}\} +
+\gamma_{3} \{\text{Win percent last year}\} +
+\gamma_{4} \{\text{Preseason odds}\} +
+\gamma_{5} \{\text{Preseason payroll}\} +
+\epsilon_{ij}
+\; $$
+
+### Level 3: Predicted Google hits by month
+
+## Results
+
+(A few plots and discussion of how well the model fits)
+
+## Measuring bandwagoning
 

### 2020-09-01 Paper_Outline.Rmd
added more to paper

--- 2020-08-29 Paper_Outline.Rmd
@@
         panel.grid.minor.x = element_blank(),
+        panel.grid.major.y = element_line(colour = "gray92"),
+        panel.grid.minor.y = element_line(colour = "gray92"),
         axis.title.x = element_blank(),
         axis.title.y = element_text(angle = 0, vjust = 0.5),
-        panel.border = element_rect(colour = "gray85", fill = NA))
+        panel.border = element_rect(colour = "gray80", fill = NA))
 ```
@@
         panel.grid.minor.x = element_blank(),
+        panel.grid.major.y = element_line(colour = "gray92"),
+        panel.grid.minor.y = element_line(colour = "gray92"),
         axis.title.y = element_text(angle = 0, vjust = 0.5),
         axis.title.x = element_blank(),
-        panel.border = element_rect(colour = "gray85", fill = NA))
+        panel.border = element_rect(colour = "gray80", fill = NA))
 
@@
         panel.grid.minor = element_blank(),
-        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.border = element_rect(colour = "gray80", fill = NA),
         panel.background = element_rect(fill = "white")
@@
 
-As seen in Figure 3a and 3b, the skew is too extreme to address with any common data 
+As seen in Figures 3a and 3b, the skew is too extreme to address with any common data 
 transformations.  However, we assert that the raw data values are not meaningful in 
@@
   labs(title = "Quantile-Normalized Google Trend Hits for SF Giants (Past 5 Years)") + 
-  theme(plot.title = element_text(face = "bold", size = 13),
+  theme(
         axis.title.x = element_blank(),
@@
         panel.grid.minor.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "gray85"),
-        panel.grid.minor.y = element_line(colour = "gray85"),
-        panel.border = element_rect(colour = "gray85", fill = NA),
-        panel.background = element_rect(fill = "white"),
-        plot.background = element_rect(fill = "darkorange"))
+        panel.grid.major.y = element_line(colour = "gray92"),
+        panel.grid.minor.y = element_line(colour = "gray92"),
+        panel.border = element_rect(colour = "gray80", fill = NA),
+        panel.background = element_rect(fill = "white")
+  )
 
@@
   annotate(geom = "text", x=11.5, y=6.5, label = "Off-Season") +
-  labs(title = "Monthly Average of Google Trend Hits for SF Giants (2004-2020)") + 
+  labs(title = "Monthly Average Google Trend Hits for SF Giants (2004-2020)") + 
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
-  scale_y_discrete(expand = c(0,0), limits = factor(0:9)) +
+  scale_y_continuous(expand = c(0,0), limits = c(0,10)) + 
   theme(
-      # plot.title = element_text(face = "bold"),
         axis.title.x = element_blank(),
@@
         panel.grid.major.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "gray90"),
+        panel.grid.major.y = element_line(colour = "gray92"),
         panel.grid.minor.y = element_blank(),
-        panel.border = element_rect(colour = "gray75", fill = NA),
+        panel.border = element_rect(colour = "gray80", fill = NA),
         panel.background = element_rect(fill = "white")
-        # plot.background = element_rect(fill = "darkorange")
   )
@@
   labs(title = "Quantile-Normalized Google Trend Hits for SEA Mariners (Past 5 Years)") + 
-  theme(plot.title = element_text(face = "bold", size = 13),
+  theme(
         axis.title.x = element_blank(),
@@
         panel.grid.minor.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "gray85"),
-        panel.grid.minor.y = element_line(colour = "gray85"),
-        panel.border = element_rect(colour = "gray85", fill = NA),
-        panel.background = element_rect(fill = "white"),
-        plot.background = element_rect(fill = "cyan4"))
+        panel.grid.major.y = element_line(colour = "gray92"),
+        panel.grid.minor.y = element_line(colour = "gray92"),
+        panel.border = element_rect(colour = "gray80", fill = NA),
+        panel.background = element_rect(fill = "white")
+        )
 
@@
   annotate(geom = "text", x=11.5, y=1.75, label = "Off-Season") +
-  labs(title = "Monthly Average of Google Trend Hits for SEA Mariners (2004-2020)") + 
+  labs(title = "Monthly Average Google Trend Hits for SEA Mariners (2004-2020)") + 
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
-  scale_y_continuous(expand = c(0,0), limits = c(0, 3.5)) +
+  scale_y_continuous(expand = c(0,0), limits = c(0,4)) +
   theme(
-        # plot.title = element_text(face = "bold"),
         axis.title.x = element_blank(),
@@
         panel.grid.major.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "gray90"),
-        panel.grid.minor.y = element_line(colour = "gray90"),
+        panel.grid.major.y = element_line(colour = "gray92"),
+        panel.grid.minor.y = element_blank(),
         panel.border = element_rect(colour = "gray80", fill = NA),
         panel.background = element_rect(fill = "white")
-        # plot.background = element_rect(fill = "cyan4")
         )
@@
   labs(title = "Average Quantile-Normalized Google Trend Hits for MLB Teams (Past 5 Years)") + 
-  theme(plot.title = element_text(face = "bold", size = 13),
+  theme(
         axis.title.x = element_blank(),
@@
         panel.grid.minor.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "gray85"),
-        panel.grid.minor.y = element_line(colour = "gray85"),
-        panel.border = element_rect(colour = "gray85", fill = NA),
-        panel.background = element_rect(fill = "white"),
-        plot.background = element_rect(fill = "azure3"))
+        panel.grid.major.y = element_line(colour = "gray92"),
+        panel.grid.minor.y = element_line(colour = "gray92"),
+        panel.border = element_rect(colour = "gray80", fill = NA),
+        panel.background = element_rect(fill = "white")
+        )
 
@@
   annotate(geom = "text", x=11.5, y=2.5, label = "Off-Season") +
-  labs(title = "Monthly Average of Google Trend Hits for MLB Teams (2004-2020)") + 
+  labs(title = "Monthly Average Google Trend Hits for MLB Teams (2004-2020)") + 
   scale_x_discrete(labels = month.abb[c(1:12)]) + 
-  scale_y_continuous(expand = c(0,0), limits = c(0, 5)) +
-  theme(plot.title = element_text(face = "bold"),
+  scale_y_continuous(expand = c(0,0), limits = c(0, 5.5)) +
+  theme(
         axis.title.x = element_blank(),
@@
         panel.grid.major.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "gray85"),
+        panel.grid.major.y = element_line(colour = "gray92"),
         panel.grid.minor.y = element_blank(),
-        panel.border = element_rect(colour = "gray85", fill = NA),
+        panel.border = element_rect(colour = "gray80", fill = NA),
         panel.background = element_rect(fill = "white")
-        # plot.background = element_rect(fill = "azure3")
         )
@@
 
+MLB teams that have done well since 2004, like the San Francisco Giants, will typically have higher quantile-normalized Google trend hits over the past 5 years than the average MLB team.
+Moreover, the Giants had a higher monthly average Google hits than the average team; they even had higher spikes during the post-season because they were very successful during this time.
+On the other hand, an MLB team who is less successful, such as the Seattle Mariners, will have lower quantile-normalized Google trends over the past 5 years and a lower monthly average Google hits than the average team.
+The Mariners monthly average Google hits also severely drops during the post-season; a large contribution to this is that they have never made the World Series.
+
 ## Model-Building
+
+As mentioned, the data was quantile-normaized before construction of the model began.
 

### 2020-09-07 KB_model_2.Rmd
Updated plots more and added to paper
added more to paper
Added more to Paper

@@
 library(tidymodels)
+library(lubridate)
+library(modelr)
 ```
@@
 
+dat %>%
+  write.csv("Model_Data.csv")
 # 
@@
 
+summary(fit_overall)
+
 dat_annual <- dat_annual %>%
@@
   )
+
 ```
@@
                       data = dat_annual)
+
+summary(fit_annual)
 
@@
                       data = dat)
+
+summary(fit_monthly)
 
@@
 ```{r}
-lm(resid_final ~ team:phase, data = dat)
+test <- lm(resid_final ~ team:phase, data = dat)
 ```
